---
title: "HW 1 Solutions"
date: "9/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("bayesplot","knitr","arm","ggplot2","rstanarm")
```

## 7.2 Fake-data simulation and regression: 
Simulate 100 data points from the linear model, y =  a + bx + error, with a = 5, b = 7, the values of x being sampled at random from a uniform distribution on the range [0, 50], and errors that are normally distributed with mean 0 and standard deviation 3. 

### 7.2a 
Fit a regression line to these data and display the output. 

```{r}
set.seed(2022)
a <- 5
b <- 7
x <- runif(100,0,50)
error <- rnorm(100,0,3)
y <- a+b*x+error
model <- lm(y~x)
summary(model)
```

### 7.2b 
Graph a scatterplot of the data and the regression line. 

```{r}
ggplot(model) + aes(x = x, y = y) + 
  geom_point(alpha = 0.3) + ylab("y") + 
  xlab("x") +
  geom_smooth(method='lm',formula= y~x)
```

### 7.2c 
Use the text function in R to add the formula of the fitted line to the graph. 

```{r}
ggplot(model) + aes(x = x, y = y) + 
  geom_point(alpha = 0.3) + ylab("y") + 
  xlab("x") +
  geom_smooth(method='lm',formula= y~x) +
  geom_text(x=10,y=150,label = "y=5.309+7.010x+error")
```

## 7.3 Fake-data simulation and fitting the wrong model: 
Simulate 100 data points from the model,  y = a + bx + cx2 + error, with the values of x being sampled at random from a uniform  distribution on the range [0, 50], errors that are normally distributed with mean 0 and standard  deviation 3, and a, b, c chosen so that a scatterplot of the data shows a clear nonlinear curve. 


### 7.3 a
Fit a regression line stan_glm(y ~ x) to these data and display the output. 

```{r}
a <- 5
b <- 7
c <- 9
x <- runif(100,0,50)
error <- rnorm(100,0,3)
y <- a+b*x+c*(x^2)+error
df <- data.frame(x,y)
stan_model <- stan_glm(y ~ x, data=df, refresh=0)
stan_model
```

### 7.3b
Graph a scatterplot of the data and the regression line. This is the best-fit linear regression.  What does “best-fit” mean in this context?

```{r}
plot(x, y, xlab="x", ylab="y")
abline(coef(stan_model), col="purple")
# The best-fit refers to "best approximation" of data points in the regression.
```


## 7.6 Formulating comparisons as regression models: 
Take the election forecasting model and simplify it by creating a binary predictor defined as x = 0 if income growth is less than 2% and x = 1 if income growth is more than 2%. 

```{r}
library(rosdata)
hibbs$x[hibbs$growth > 2] <- 1
hibbs$x[hibbs$growth < 2] <- 0
```

### 7.6a
Compute the difference in incumbent party’s vote share on average, comparing those two groups of elections, and determine the standard error for this difference.

```{r}
avg_share <- aggregate(hibbs$vote, list(hibbs$inc_party_candidate), FUN=mean)
avg_share
# I want to compare Carter and Bush Jr.
Carter <- avg_share[avg_share$Group.1 == "Carter", "x"]
B_Jr <- avg_share[avg_share$Group.1 == "Bush, Jr.", "x"]
B_Jr - Carter
```

### 7.6b
Regress incumbent party’s vote share on the binary predictor of income growth and check that the resulting estimate and standard error are the same as above. 

```{r}
b_model <- lm(hibbs$vote ~ hibbs$x)
summary(b_model)
```

## 8.8 Comparing lm and stan_glm: 
Use simulated data to compare least squares estimation to default Bayesian regression: 

### 8.8a
Simulate 100 data points from the model, y = 2 + 3x + error, with predictors x drawn from a uniform distribution from 0 to 20, and with independent errors drawn from the normal distribution with mean 0 and standard deviation 5. Fit the regression of y on x data using  lm and stan_glm (using its default settings) and check that the two programs give nearly identical results. 

```{r}
set.seed(2022)
a <- 2
b <- 3
x <- runif(100,0,20)
error <- rnorm(100,0,5)
y <- a+b*x+error
model <- lm(y~x)
summary(model)
stan_model <- stan_glm(y~x,refresh=0)
summary(stan_model)
```

### 8.8b
Plot the simulated data and the two fitted regression lines. 

```{r}
ggplot(model) + aes(x = x, y = y) + 
  geom_point(alpha = 0.3) + ylab("y") + 
  xlab("x") +
  geom_smooth(method='lm',formula= y~x)

plot(x, y, xlab="x", ylab="y")
abline(coef(stan_model), col="purple")
```

### 8.8c
Repeat the two steps above, but try to create conditions for your simulation so that lm and  stan_glm give much different results. 

```{r}
```

## 10.1 Regression with interactions: 
Simulate 100 data points from the model, y = b0 + b1 x +  b2 z + b3 xz + error, with a continuous predictor x and a binary predictor z, coefficients  b = c(1, 2, -1, -2), and errors drawn independently from a normal distribution with mean 0  and standard deviation 3, as follows. For each data point i, first draw zi, equally likely to take on the values 0 and 1. Then draw xi from a normal distribution with mean zi and standard deviation 1. Then draw the error from its normal distribution and compute yi. 

### 10.1a
Display your simulated data as a graph of y vs. x, using dots and circles for the points with z = 0 and 1, respectively. 

```{r}
set.seed(2022)
a <- 2
b0 <- 1
b1 <- 2
b2 <- -1
b3 <- -2
error <- rnorm(100,0,3)
z <- c(rep(0,50),rep(1,50))
x <- rnorm(100,0.5,1)
y <- b0 + b1*x + b2*z + b3*x*z + error
df <- data.frame(x,y,z)
df$z[df$z == 0] <- "Group 0"
df$z[df$z == 1] <- "Group 1"

ggplot(df, aes(x=x, y=y, group=z)) +
  geom_point(aes(shape=z), size=4) +
  scale_shape_manual(values=c(16,1)) + 
  theme(legend.position="top")
```

### 10.1b
Fit a regression predicting y from x and z with no interaction. Make a graph with the data and two parallel lines showing the fitted model. 

```{r}
y <- b0 + b1*x + b2*z + b3*x*z + error
df <- data.frame(x,y,z)
stan_model <- stan_glm(y ~ x + z,data=df,refresh=0)
colors <- ifelse(df$z == 1, "red", "dark green")
plot(df$x, df$y,
  xlab="x", ylab="y", col=colors, pch=20)
b_hat <- coef(stan_model)
abline(b_hat[1] + b_hat[2], b_hat[3], col="red")
abline(b_hat[1], b_hat[3], col="dark green")
```

### 10.1c
Fit a regression predicting y from x, z, and their interaction. Make a graph with the data  and two lines showing the fitted model. 

```{r}
int_model <- stan_glm(y ~ x + z + x*z, data=df,refresh=0)
colors <- ifelse(df$z==1, "orange", "light blue")
plot(df$x, df$y,
  xlab="x", ylab="y", col=colors, pch=20)
b_hat <- coef(int_model)
abline(b_hat[1] + b_hat[2], b_hat[3] + b_hat[4], col="orange")
abline(b_hat[1], b_hat[3], col="light blue")
```


## 10.2 Regression with interactions: 
Here is the output from a fitted linear regression of outcome y on pre-treatment predictor x, treatment indicator z, and their interaction: 

```{r}
```

### 10.2a
Write the equation of the estimated regression line of y on x for the treatment group and the  control group, and the equation of the estimated regression line of y on x for the control group. 

```{r}
# Treatment & Control group:
# y = 1.2 + 1.6x + 2.7z
```

### 10.2b
Graph with pen on paper the two regression lines, assuming the values of x fall in the range  (0, 10). On this graph also include a scatterplot of data (using open circles for treated units  and dots for controls) that are consistent with the fitted model. 

```{r}
```

## 10.5 Regression modeling and prediction: 
The folder KidIQ contains a subset of the children and mother data discussed earlier in the chapter. You have access to children’s test scores at age 3,  mother’s education, and the mother’s age at the time she gave birth for a sample of 400 children. 

```{r}
data("kidiq")
head(kidiq)
```

### 10.5a
Fit a regression of child test scores on mother’s age, display the data and fitted model,  check assumptions, and interpret the slope coefficient. Based on this analysis, when  do you recommend mothers should give birth? What are you assuming in making this recommendation? 

```{r}
age_model <- lm(kid_score ~ mom_age, data=kidiq)
plot(kidiq$mom_age, kidiq$kid_score, xlab="Mother Age", ylab="Child test score")
abline(coef(age_model))
print(age_model)
# The intercept is 70.9569. If mom is 0-year-old, her kid is predicted to have a test score of 70.9569.

# The mom_age coefficient(slope) is 0.6952. With every 1 year of increase in mom's age when giving birth, kid's score is predicted to increase by 0.6952.

# In this data, mom's age varies from 17 to 29. Based on the plot, we have a positive relationship between mom's age and kid's test score. The kid with highest test score whose mom is 25 years old, and the second highest is 23. I would suggest mothers give birth between 23 and 29 years old.
```

### 10.5b
Repeat this for a regression that further includes mother’s education, interpreting both slope  coefficients in this model. Have your conclusions about the timing of birth changed? 

```{r}
education_model <- lm(kid_score ~ mom_age + mom_hs, data=kidiq)
summary(education_model)
colors <- ifelse(kidiq$mom_hs==1, "pink", "blue")
plot(kidiq$mom_age, kidiq$kid_score,
  xlab="Mother Age", ylab="Child test score", col=colors, pch=20)
b_hat <- coef(education_model)
# The intercept is 70.4787. If mom is 0-year-old and did not graduate from high school, 
# her kid is predicted to have a test score of 70.4787.
# The coefficient of mom_age is 0.3261. With every 1 year of increase in mom's age when 
# giving birth, kid's score is predicted to increase by 0.3261. 
# The coefficient of mom_hs is 11.3112. If two mothers give birth at the same age, the 
# kid is expected to have 11.3112 points higher than another kid whose mom did not 
# graduate from high school.
# My conclusion did not change because it supports that mothers give birth after high 
# school education.
```

### 10.5c
Now create an indicator variable reflecting whether the mother has completed high school or not. Consider interactions between high school completion and mother’s age. Also create a plot that shows the separate regression lines for each high school completion status group. 

```{r}
stan_model <- stan_glm(kid_score ~ mom_hs + mom_age + mom_hs:mom_age, data=kidiq,
                  refresh = 0)
colors <- ifelse(kidiq$mom_hs==1, "pink", "blue")
plot(kidiq$mom_age, kidiq$kid_score,
  xlab="Mother Age", ylab="Child test score", col=colors, pch=20)
b_hat <- coef(stan_model)
abline(b_hat[1] + b_hat[2], b_hat[3] + b_hat[4], col="pink")
abline(b_hat[1], b_hat[3], col="blue")
```

### 10.5d
Finally, fit a regression of child test scores on mother’s age and education level for the first  200 children and use this model to predict test scores for the next 200. Graphically display  comparisons of the predicted and actual scores for the final 200 children. 

```{r}
kid_head <- kidiq[1:200,]
kidhead_model <- lm(kid_score ~ mom_age+mom_hs, data=kid_head)
summary(kidhead_model)
# Based on the regression with first 200 kids, the regression model is kid_score = 82.9869+0.2988*mom_age+3.6596*mom_hs.
kidhead_plot <- plot(kid_head$mom_age, kid_head$kid_score,
  xlab="Mother Age", ylab="Child test score", col=colors, pch=20)

kid_tail <- tail(kidiq,200)
kidtail_model <- lm(kid_score ~ mom_age+mom_hs, data=kid_tail)
kidtail_plot <- plot(kid_tail$mom_age, kid_tail$kid_score,
  xlab="Mother Age", ylab="Child test score", col=colors, pch=20)
```

## 10.6 Regression models with interactions: 
The folder Beauty contains data (use file beauty.csv)  Beauty and  teaching evaluations from Hamermesh and Parker (2005) on student evaluations of instructors’ beauty and teaching quality for several courses at the University of Texas. The teaching evaluations were conducted at the end of the semester, and the beauty judgments were made later, by six students who had not attended the classes and were not aware of the course evaluations. 

See also Felton, Mitchell, and Stinson (2003) for more on this topic. 

```{r}

```


### 10.6a
Run a regression using beauty (the variable beauty) to predict course evaluations (eval), adjusting for various other predictors. Graph the data and fitted model, and explain the meaning of each of the coefficients along with the residual standard deviation. Plot the residuals versus fitted values. 

```{r}
beauty_model <- lm(eval ~ beauty,beauty)
summary(beauty_model)
# The intercept is 4.01002. If the instructor's beauty is evaluated as zero, The course 
# evaluation is predicted to be 4.01002.
# The beauty coefficient is 0.13300. With every 1 point increase in beauty, the course 
# evaluation is predicted to increase by 0.13300.

plot(beauty$beauty, beauty$eval, xlab="Beauty", ylab="Average teaching evaluation")
coefs <- coef(beauty_model)
curve(coefs[1] + coefs[2]*x, add=TRUE)

res <- resid(beauty_model)
plot(fitted(beauty_model), res)
abline(0,0)
```

### 10.6b
Fit some other models, including beauty and also other predictors. Consider at least one model with interactions. For each model, explain the meaning of each of its estimated coefficients.

```{r}
head(beauty)
minor_model <- lm(eval ~ beauty + minority + beauty*minority, data=beauty)
summary(minor_model)
# The intercept is 4.031672. If the instructor's beauty is evaluated as zero and 
# is not from minority groups, the course evaluation is predicted to be 4.03167.

# The beauty coefficient is 0.16584. It can be thought of as the comparison 
# of mean course evaluation scores across instructors who are not minorities,
# but whose beauty scores differ by 0.16584 point.

# The minority coefficient is -0.14141. If instructor from minority group:eval = 3.89026-0.08273*beauty. With 1 year old increase in age, the beauty is predicted 
# to decrease by 0.08273.

# If instructor not from minority group:eval = 4.03167+0.16584*beauty. With 1 year 
# old increase in age, the beauty is predicted to increase by 0.16584.

# The interaction coefficient is -0.24857. The increase of effectiveness of beauty 
# for a 1 unit change in If instructor from minority group:
# eval = 3.89026-0.08273*beauty. If instructor not from minority group:eval = 4.03167+0.16584*beauty.

```

## 10.7 Predictive simulation for linear regression:
Take one of the models from the previous exercise.

### 10.7a
Instructor A is a 50-year-old woman who is a native English speaker and has a beauty score  of -1. Instructor B is a 60-year-old man who is a native English speaker and has a beauty  score of -0.5. Simulate 1000 random draws of the course evaluation rating of these two instructors. In your simulation, use posterior_predict to account for the uncertainty in  the regression parameters as well as predictive uncertainty. 

```{r}

```

### 10.7b
Make a histogram of the difference between the course evaluations for A and B. What is the  probability that A will have a higher evaluation? 

```{r}
```

## 10.8 How many simulation draws: 
Take the model from Exercise 10.6 that predicts course evaluations from beauty and other predictors. 

### 10.8a
Display and discuss the fitted model. Focus on the estimate and standard error for the coefficient of beauty. 

```{r}
beauty_model <- stan_glm(eval ~ beauty,data=beauty,refresh=0)
summary(beauty_model)
```

### 10.8b
Compute the median and mad sd of the posterior simulations of the coefficient of beauty,and check that these are the same as the output from printing the fit. 

```{r}
sims <- as.matrix(beauty_model)
Median <- apply(sims, 2, median) 
MAD_SD <- apply(sims, 2, mad) 
print(cbind(Median, MAD_SD))
```

### 10.8c
Fit again, this time setting iter = 1000 in your stan_glm call. Do this a few times in order  to get a sense of the simulation variability. 

```{r}
```

### 10.8d
Repeat the previous step, setting iter = 100 and then iter = 10. 

```{r}
```

### 10.8e
How many simulations were needed to give a good approximation to the mean and standard  error for the coefficient of beauty? 




